Vibronic Coupling Effects in the Photoelectron Spectrum of Ozone: A Coupled-Cluster Approach

### 2024-08-17 article.tex
Split Results into Results and Discussion

\documentclass[12pt,pra,aps,superscriptaddress]{revtex4-2}
\usepackage{amsmath,amssymb,physics}
\usepackage{hyperref} % Clickable links in the document
\usepackage{graphicx} % For figures with images
\usepackage{natbib} % bibliography
\setcitestyle{super, sort&compress, comma} % chemistry style
\usepackage{hyperref} % Clickable links in the document and their colors
\hypersetup{
    colorlinks=true, % false: boxed links; true: colored links
    linkcolor=cyan, % color of internal links
    citecolor=magenta, % color of links to bibliography
    filecolor=magenta, % color of file links
    urlcolor=cyan, % color of external links
    runcolor=cyan
}

@@

\author{John F. Stanton}
\email[Electronic address: ]{jfstanton@gmail.com}
\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

@@

\section{Methods}

@@

Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
experimental one taken from reference~\cite{dyke:O3:74}. A comparison of the
simulated and experimental spectra is presented on
Figure~\ref{fig:sim_vs_dyke} and shows a good match. Peak A is known to be a
hot band.~\cite{KDC:O3:92} The simulation reproduces well the consecutive
increase in the intensities of peaks B, C, D, and E. The spacing between these
peaks is also well reproduced. Drop in the intensity of peak F is also
captured by the simulation. Starting from peak G the simulation shows
discrepancy with the experiment. A sudden drop in the intensity past the peak
H is not observed in the simulation. We discuss a likely source of this
mismatch later.

\begin{figure}
    \includegraphics[width=10 cm]{figures/spectrum_overline}
    \caption{
        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
        scale in eV. Top axis shows energy offset from the origin in
        cm$^{-1}$. Stick spectrum shows positions and intensities of all
        simulated states. Blue color corresponds to the oscillator strength
        originating from the $^2$A$_1$ basis state while the red ones
        correspond to $^2$B$_2$. Gray vertical lines with captions on top
        indicate positions of features as measured by the PFI-ZEKE
        experiment.~\cite{Willitsch:O3ZEKE:2005} 
        $D_0$ marks the dissociation threshold of O$_3^+$.
        Gray dotted line marks the energy of the minimum of the conical
        intersection (CI).
    }
    \label{fig:ozone_overlay}
\end{figure}

Our simulation allows for an additional element of analysis of the simulated
spectrum. Figure~\ref{fig:ozone_overlay} presents the same simulated spectrum
as in Figure~\ref{fig:sim_vs_dyke} with additional information on the
decomposition of the spectrum.  All lines that contribute to the spectrum are
marked individually and in two different colors. The spectrum shows that peaks
A and B are almost purely of the $^2$A$_1$ character. Starting from peak C
each the contributions from two states are equally important, it is also clear
that each peak of the low-resolution spectrum has in fact contributions from
many vibronic peaks. At the energy of about $2000$~cm$^{-1}$ above the origin
the density of vibronic peaks increases significantly. This value can be
compared to the minimum of the conical intersection, which our simulation
locates at $3174$~cm$^{-1}$ above the origin ($12.92$~eV). Our simulation is
incapable of accounting for dissociation of the molecule, therefore the we
expect a discrepancy with the experiment as the energy gets closer to the
dissociation threshold located at $4840$~cm$^{-1}$ above the
origin.~\cite{Willitsch:O3ZEKE:2005}

@@

We would like to assign the vibronic peaks from our simulation. To this end,
we run the simulation once again, this time, however we set the linear
diabatic constant to zero, i.e., $\lambda = 0$. With that change we can
reproduce the spectrum (at an equivalent level of theory) but in an artificial
case where there is no vibronic coupling. Figure~\ref{fig:no_coupling}
presents this spectrum. The non-coupled spectrum is easy to assign using the
labels that mark the symmetry of the electronic state, A$_1$ or B$_2$, and the
vibrational state ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of
quanta in mode $i$ with $i=1$ for the symmetric stretch, $i=2$ for the
symmetric bend and $i=3$ for the asymmetric stretch. The assigned spectrum
shows progressions in the symmetric bend. There is one such progression in
each electronic states.  Additionally for each state, there is also another
progression in the symmetric bend with one excitation in the symmetric
stretch.

@@

We decompose the vibronic states of our main simulation in the basis of the
artificial, uncoupled states discussed in the previous paragraph.
Figure~\ref{fig:sim_vs_zeke} shows the assigned spectrum and
Table~\ref{tab:peak_assignment} lists the decomposition of all peaks in the
region of up to $3000$~cm$^{-1}$ with intensities larger than $10^{-3}$. We
compare the assigned spectrum against the high-resolution
pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
2005.\cite{Willitsch:O3ZEKE:2005} The PFI-ZEKE spectrum is the best source of
information on the location of the peaks, especially the origin, against which
align our simulation. The origin of the simulated spectrum is located
$170$~cm$^{-1}$ lower than the experimental origin which was experimentally
observed at $101,020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
our error estimate for the vertical ionization energy
($30$~meV$=240$~cm$^{-1}$).  Once the origin locations are aligned the
comparison shows a good match between the simulation and the experiment.
Especially the states with an oscillator strength originating from the A$_1$
state are well aligned with the experimental features.

States close to the origin of the band are similar to the non-coupled states.
The origin peak is of a clear A$_1(0,0,0)$ character, the first two
excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are also very
similar to the non-coupled states, while the higher excitations in this
progression are showing large mixing. The same progression with one
vibrational quantum in the symmetric stretch is more interesting. The first of
its peaks is very well aligned with an experimentally visible feature, which
was previously assigned as the origin of the second state. The second peak in
this series, the first combination state A$_1(1,1,0)$, has very high
similarity to its uncoupled version. It is a good candidate to assign the
experimental, unassigned feature above $102,600$~cm$^{-1}$.

States with an oscillator strength from the B$_2$ state are on the other hand
not aligning well with the experiment. First such peak, close to the $12.64$~eV
mark on the Figure~\ref{fig:sim_vs_zeke}, is of a vibronic character and we
assign it as a A$_1$(0,0,1) state which gains intensity thanks to the coupling
to the B$_2$ state. This vibronic peak lies in the part of the spectrum marked
as pollution due to water. The origin of the B$_2$ state is very similar to
the uncoupled B$_2(0,0,0)$ state, but it is located in an empty area of the
experimental spectrum. The next peak, slightly above the $12.76$~eV mark on
Figure~\ref{fig:sim_vs_zeke}, corresponds to the one excitation of the symmetric
bend in the B$_2$ state, but as the label on the figure shows, it is only
about $55$\% similar to the uncoupled state.

@@

Finally, we compare our simulation to an earlier accurate simulation, the work
of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011} Both spectra show a
good match to the experiment and agree in the assignment of the first
progression in the bending mode. The previous simulation, however, shows
smaller spacing between other lines, showing much higher congestion of the
spectrum. Additionally the origin of the second state falls at lower energy
aligning well with the peak that was assigned in the PFI-ZEKE also as the
origin of the second state.  While an additional simulation of the line
intensities would allow for the most complete comparison to the experimental
spectrum, the strong alignment of the simulated peaks of a leading
A$_1$ character with the experimental features leads us to believe that our
simulation offers the best assignment of the spectrum to date.

@@

@@

This project was initiated when three of the authors (P.W, P.G.S. and J.F.S)
were in Budapest, where J.F.S. was serving as a recipient of the John von
Neumann Award in STEM bestowed by the Fulbright Foundation. Additional
research presented here benefited from the NSF Center for Chemical Innovation
Phase I (grant no. CHE-2221453) and U.S. Department of Energy, Basic Energy
Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
Prof. Bartlett a happy ninth decade of life, and hope that he continues to
stimulate others in the field with his creative insights.

@@

\bibliographystyle{prf}
\bibliography{allrefs}

>>>>

\documentclass[12pt,a4paper,prb,aps,superscriptaddress]{revtex4}
\usepackage{amsmath,amssymb,physics}
\usepackage{graphicx} % For figures with images
\usepackage{natbib} % bibliography
\setcitestyle{super, sort&compress, comma} % chemistry style
\usepackage{hyperref} % Clickable links in the document and their colors
\hypersetup{
    colorlinks=true, % false: boxed links; true: colored links
    linkcolor=cyan, % color of internal links
    citecolor=magenta, % color of links to bibliography
    filecolor=magenta, % color of file links
    urlcolor=cyan, % color of external links
    runcolor=cyan
}

@@

\author{John F. Stanton}
\email[Electronic address: ]{jfstanton137@gmail.com}
\affiliation{Quantum Theory Project, Departments of Chemistry and Physics, University of Florida, Gainesville, FL, USA 32611}
\affiliation{Laboratory of Theoretical Chemistry, Institute of Chemistry, ELTE
E{\"o}tv{\"o}s Lor{\'a}nd University, P{\'a}zm{\'a}ny P{\'e}ter stny. 1/A,
Budapest, Hungary}

@@

\section{Theory}

@@

Figure~\ref{fig:sim_vs_dyke} compares the simulated spectrum to the
experimental one taken from reference~\cite{dyke:O3:74}. Our simulation allows
for an additional element of analysis of the simulated spectrum.
Figure~\ref{fig:ozone_overlay} presents a decomposition of the simulated
spectrum from Figure~\ref{fig:sim_vs_dyke}. All lines that contribute to the
spectrum are marked individually and are color-coded indicating which
electronic state's transition intensity the peak draws from. The total
envelope of the spectrum is also decomposed showing contributions from both
states. Our simulation locates the minimum of the conical intersection at
$3174$~cm$^{-1}$ above the origin ($12.92$~eV) that is marked on the figure as
CI.

\begin{figure}
    \includegraphics[width=10 cm]{figures/spectrum_overline}
    \caption{
        Simulated photoelectron spectrum of ozone. Bottom axis shows energy
        scale in eV. Top axis shows energy offset from the origin in
        cm$^{-1}$. Stick spectrum shows positions and intensities of all
        simulated states. Blue color corresponds to the oscillator strength
        originating from the $^2$A$_1$ basis state while the red one
        corresponds to $^2$B$_2$. Gray vertical lines with captions on top
        indicate positions of features as measured by the PFI-ZEKE
        experiment.~\cite{Willitsch:O3ZEKE:2005} The simulated spectrum was
        shifted to match the PFI-ZEKE experimental origin. $D_0$ marks the
        dissociation threshold of O$_3^+$. Gray dotted line marks the energy
        of the minimum of the conical intersection (CI).
    }
    \label{fig:ozone_overlay}
\end{figure}

@@

We would like to assign the vibronic peaks from our simulation. To this end,
we run the simulation once again, this time, however we set the linear
diabatic constant to zero, i.e., $\lambda = 0$. With that change we can
reproduce the spectrum (at an equivalent level of theory) but in an artificial
case where there is no vibronic coupling. Figure~\ref{fig:no_coupling}
presents this spectrum. The non-coupled spectrum is easy to assign using the
labels that mark the symmetry of the electronic state, A$_1$ or B$_2$, and the
vibrational state ($\nu _1 \nu_2 \nu_3$), where $\nu _i$ is the number of
quanta in mode $i$ with $i=1$ for the symmetric stretch, $i=2$ for the
symmetric bend and $i=3$ for the asymmetric stretch. The assigned spectrum
shows progressions in the symmetric bend. There is one such progression in
each electronic state.  Additionally for each state, there is also another
progression in the symmetric bend with one excitation in the symmetric
stretch.

@@

We decompose the vibronic states of our main simulation in the basis of the
artificial, uncoupled states discussed in the previous paragraph.
Figure~\ref{fig:sim_vs_zeke} shows the assigned spectrum and
Table~\ref{tab:peak_assignment} lists the decomposition of all peaks in the
region of up to $3000$~cm$^{-1}$ with intensities larger than $10^{-3}$. We
compare the assigned spectrum against the high-resolution
pulsed-field-ionization zero-kinetic-energy (PFI-ZEKE) spectrum from
2005.\cite{Willitsch:O3ZEKE:2005} The PFI-ZEKE spectrum is the best source of
information on the location of the peaks, especially the origin, against which
align our simulation. The origin of the simulated spectrum is located
$170$~cm$^{-1}$ lower than the experimental origin which was experimentally
observed at $101,020.5$~cm$^{-1}$.\cite{Willitsch:O3ZEKE:2005} This is within
our error estimate for the vertical ionization energy
($30$~meV$=240$~cm$^{-1}$).

@@

Lastly, we compare our simulation to an earlier accurate simulation, the work
of Tarroni and Carter from 2011.~\cite{tarroni:O3:2011}
Figure~\ref{fig:sim_vs_tarronicarter} presents a comparison of lines from both
simulations. The comparison includes only the lines from the earlier
simulation that were tabulated with an assignment.

@@

\section{Discussion}

We first discuss the comparison of our simulations to the lower-resolution
experimental spectrum presented on Figure~\ref{fig:sim_vs_dyke}. The peak A is
known to be a hot band.~\cite{KDC:O3:92} The simulation reproduces well the
consecutive increase in the intensities of peaks B, C, D, and E. The spacing
between these peaks is also well reproduced. Drop in the intensity at the peak
F is captured by the simulation. Starting from the peak G, the simulation
shows discrepancy with the experiment. A sudden drop in the intensity past the
peak H is not observed in the simulation. We discuss a likely source of this
mismatch later.

The decomposition of the spectrum presented on Figure~\ref{fig:ozone_overlay}
introduces additional insight. The spectrum shows that peaks B and C are
almost purely of the $^2$A$_1$ character. Starting from the peak D, the
contributions from two states are equally important. It is also clear that the
following peaks arise as mixtures of many vibronic peaks. At the energy of
about $2000$~cm$^{-1}$ above the origin the density of vibronic peaks
increases significantly. This value can be compared to the minimum of the
conical intersection located about $1000$~cm$^{-1}$ higher. Our
simulation is incapable of accounting for dissociation of the molecule,
therefore we expect a discrepancy with the experiment as the energy gets
closer to the dissociation threshold located at $4840$~cm$^{-1}$ above the
origin.~\cite{Willitsch:O3ZEKE:2005}

Moving on to the comparison with the high-resolution spectrum, see
Figure~\ref{fig:sim_vs_zeke}, reveals more details. Lines of the PFI-ZEKE
experiment and our simulation show a good match. Especially the states with
an oscillator strength originating from the A$_1$ state are well aligned with
the experimental features. States close to the origin are similar to the
non-coupled states. The origin peak is of a clear A$_1(0,0,0)$ character, the
first two excitations in the symmetric bend, A$_1(0,1,0)$ and A$_1(0,2,0)$ are
also very similar to the non-coupled states, while the higher excitations in
this progression are showing large mixing. The same progression with one
vibrational quantum in the symmetric stretch is more interesting. The first of
its peaks is very well aligned with an experimentally visible feature, which
was previously assigned as the origin of the second state. The second peak in
this series, the first combination state A$_1(1,1,0)$, has very high
similarity to its uncoupled version. It is a good candidate to assign the
experimental, unassigned feature above $102,600$~cm$^{-1}$.

States with an oscillator strength from the B$_2$ state are, on the other
hand, not aligning well with the experiment. First such peak, close to the
$12.64$~eV mark on Figure~\ref{fig:sim_vs_zeke}, is of a vibronic
character and we assign it as a A$_1$(0,0,1) state which gains intensity
thanks to the coupling to the B$_2$ state. This vibronic peak lies in the part
of the spectrum marked as pollution due to water. The origin of the B$_2$
state is very similar to the uncoupled B$_2(0,0,0)$ state, but it is located
in an empty area of the experimental spectrum. The next peak, slightly above
the $12.76$~eV mark on Figure~\ref{fig:sim_vs_zeke}, corresponds to the one
excitation of the symmetric bend in the B$_2$ state, but as the label on the
figure shows, it is only about $55$\% similar to the uncoupled state.

The comparison to the earlier simulation on
Figure~\ref{fig:sim_vs_tarronicarter} shows that both present a good match to
the experiment. Both simulations also agree in the assignment of the first
progression in the bending mode. The previous simulation, however, shows
smaller spacing between the remaining lines, showing much higher congestion of
the spectrum. Additionally the origin of the second state falls at lower
energy aligning well with the peak that was assigned in the PFI-ZEKE
experiment also as the origin of the second state. While an additional
simulation of the line intensities would allow for the most complete
comparison to the experimental spectrum, the strong alignment of the simulated
peaks of a leading A$_1$ character with the experimental features leads us to
believe that our simulation offers the best assignment of the spectrum to
date.

@@

This project was initiated when three of the authors (P.W, P.G.S. and J.F.S)
were in Budapest, where J.F.S. was serving as a recipient of the John von
Neumann Award in STEM bestowed by the Fulbright Foundation. Additional
research presented here benefited from the NSF Center for Chemical Innovation
Phase I (grant no. CHE-2221453) and the U.S. Department of Energy, Basic Energy
Sciences (grant no. DE-FG02-05ER15629).  All the authors of this research wish
Prof. Bartlett a happy ninth decade of life, and hope that he continues to
stimulate others in the field with his creative insights.

@@

\bibliography{allrefs}
